TBD

### 2019-02-06 inverse_problems/inverse_problems.tex
Revert "updates on the Poisson process"
This reverts commit 81fde221912c2f7f1333bc0ca20606ddec02e8df.

Until now we have presented models assuming that the signals' occurrences are well separated. This condition can be alleviated by assuming a Poisson distribution on $s$.

Let $X\in\RL$ be a random vector drawn from some fixed distribution. 
We observe a random vector $Y\in\R^N$ generated as follows. Points are chosen in $\{0,\dots,N-L\}$ according to a Poisson process with parameter $\gamma N$. 
The parameter $\gamma$ plays the role of the density of the signal in the measurement. 
For each point $i$ that is chosen from $1$ to $N-L$, a random vector $X$ from the distribution is then placed in the large vector, with element $0$ at location $i$, with overlapping vectors being added together. 
If the distribution of $X$ is a single Dirac function, then we observe multiple repetition of the same signals, like in the homogeneous case. If the distribution consists of $K$ Diracs, then it is a heterogeneous model.

If $M_i$ denotes the number of hits at location $i$, $1 \le i \le N$, then by definition of the Poisson process $M_i$'s are i.i.d.\ and $M_i \sim \Poisson(\gamma)$. Conditional on the value of $M = (M_1,\dots,M_N)$, if we let $X_1^{i},\dots,X_{M_i}^i$ denote the random vectors with position 0 located at $i$, then $X_{k_1}^{i}$ and $X_{k_2}^{i}$ are independent for $k_1 \ne k_2$.
With this notation,  we can write each entry as:
%
\begin{align}
%
Y[i] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} X_k^{i-j}[j].
%
\end{align}

@@

\section{Autocorrelation analysis} \label{sec:AC_analysis}
\TODO{Maybe we should re-write this section for a random signal and then particularized to the homogeneous and heterogeneous cases}
For a signal $z\in \R^m$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{equation}
a_z^q[\ell_1,\ldots,\ell_{q-1}]   = \frac{1}{m} \sum_{i=-\infty}^{\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
\label{eq:ac_general}
\end{equation}
where indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
Explicitly, the first-, second- and third-order autocorrelations are given by: 
\begin{align} 
a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} z[i], \nonumber\\
a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} z[i]z[i+\ell],\\
a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2].  \nonumber \label{eq:ac_special}
\end{align}
The autocorrelation functions have symmetries. Specifically, $$a_z^2[\ell] = a_z^2[-\ell],$$ and
$$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$$
For our purposes, this will be applied both to $x$ (of length $L$) and to $y$ (of length $N$).

Under the separation condition, the relation between autocorrelations of the micrograph and those of $x$ is particularly simple, as we now show. It is useful to introduce some notation: let $M$ denote the number of occurrences of $x$ in $y$ (that is, the number of 1's in $s$), and let
\begin{equation}
\gamma  = \frac{M L}{N}
\label{eq:gamma}
\end{equation}
denote the density of $x$ in $y$ (that is, the fraction of entries of $y$ occupied by occurrences of $x$.) The separation condition %~\eqref{eq:spacing} 
imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

@@

In this section we omit the effect of the noise as it remains identical to the well-separated model.

The moments of $Y$ are defined by
%
\begin{align}
%
m^1_Y[i] = \E Y[i], \quad 0 \le i \le L-1,
%
\end{align}
%
\begin{align}
%
m^2_Y[i,j] = \E Y[i] Y[j], \quad 0 \le i,j \le L-1,
%
\end{align}
%
and
%
\begin{align}
%
m^3_Y[i,j,k] = \E Y[i] Y[j] Y[k], \quad 0 \le i,j,k \le L-1.
%
\end{align}

%We will also denote by $\L_l$ the autocorrelations of $X$, $l = 1,2,3$:
%%
%\begin{align}
%%
%\L_1 = \sum_{i=0}^{L-1} \M_1[i] \in \R,
%%
%\end{align}
%%
%\begin{align}
%%
%\L_2(\Delta) = \sum_{i=0}^{L-1} \M_2[i,i+\Delta]  \in \R^L,
%%
%\end{align}
%%
%and
%%
%\begin{align}
%%
%\L_3(\Delta_1,\Delta_2) = \sum_{i=0}^{L-1} \M_3[i,i+\Delta_1,i+\Delta_2]  \in \R^{L \times L}.
%%
%\end{align}
%
%Note that in the strongly-separated model, the first three observed moments are, respectively, $\L_1$, $\L_2(\Delta)$, and $\L_3(\Delta_1,\Delta_2)$. \TODO{I think we can remove it}
%%
%
We now prove that the first three autocorrelations of the well-separated model can be recovered from the moments of the Poisson process.

\begin{proposition} \label{prop:poisson}
Let $a_X^1, a_X^2$ and $a_X^3$ be the first three autocorrelation function of the well-separated model \TODO{This is a bit unclear now, but if I get it right it holds true for any distribution over $X$}. Then, for any   $0\leq i\leq L-1$ we have 
\begin{align}
m_Y^1[i] &= \gamma a_X^1,  \nonumber \\
m_Y^2[i,i+\ell] &= (\gamma a_X^1)^2 + \gamma a_X^2[\ell], \quad \ell\in[0,L-1], \\
m_Y^3[i,i+\ell_1,i+\ell_2] &= 
(\gamma a_X^1)^3 + \gamma a_X^1  \cdot ( \gamma a^2_X[\ell_1] + \gamma a_X^2[\ell_2] + \gamma a_X^2[\ell_2-\ell_1]) + \gamma a_X^3[\ell_1,\ell_2], \quad \quad \ell_1,\ell_2\in[0,L-1]. \nonumber
\end{align}
\end{proposition}

@@

%
%
%\begin{rem}
%	
%	The expected values of the non-zero (for second moment) and off-diagonal (for third moment) terms are the same as without noise, as is true for the strongly-separated case. The same proof of almost sure convergence from my notes for the strongly-separated case also goes through verbatim.
%	
%	
%\end{rem}
%

>>>>

We consider the following observation model. $X = (X[0],\dots,X[L-1])$ is a random vector of length $L$, drawn from some fixed distribution. For fixed $n$, we observe a random vector $Y$ of length $n+L$, generated as follows. Points are chosen in $\{1,\dots,n\}$ according to a Poisson process with parameter $\gamma n$. For each point $i$ that is chosen from $1$ to $n$, a random vector $X$ from the distribution is then placed in the large vector, with element $0$ at location $i$, with overlapping vectors being added together.

If $M_i$ denotes the number of hits at location $i$, $1 \le i \le n$, then by definition of the Poisson process $M_i$'s are iid and $M_i \sim \Poisson(\gamma)$. Conditional on the value of $M = (M_1,\dots,M_n)$, if we let $X_1^{i},\dots,X_{M_i}^i$ denote the random vectors with position 0 located at $i$, then $X_{k_1}^{i}$ and $X_{k_2}^{i}$ are independent for $k_1 \ne k_2$.

With this notation, if $Y \in \R^{n+L}$ is the observed vector, we can write each entry as:
%
\begin{align}
%
Y[i] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} X_k^{i-j}[j].
%
\end{align}

@@

\section{Autocorrelation analysis} \label{sec:AC_analysis}

For a signal $z\in \R^m$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{equation}
a_z^q[\ell_1,\ldots,\ell_{q-1}]   = \frac{1}{m} \sum_{i=-\infty}^{\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
\label{eq:ac_general}
\end{equation}
where indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
Explicitly, the first-, second- and third-order autocorrelations are given by: 
\begin{align} 
a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} z[i], \nonumber\\
a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} z[i]z[i+\ell],\\
a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2].  \nonumber \label{eq:ac_special}
\end{align}
The autocorrelation functions have symmetries. Specifically, $$a_z^2[\ell] = a_z^2[-\ell],$$ and
$$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$$
For our purposes, this will be applied both to $x$ (of length $L$) and to $y$ (of length $N$).

\TODO{Start directly with heterogeneous model, but first need to verify that the Poisson equivalence is true holds also for this case.}
Under the separation condition, the relation between autocorrelations of the micrograph and those of $x$ is particularly simple, as we now show. It is useful to introduce some notation: let $M$ denote the number of occurrences of $x$ in $y$ (that is, the number of 1's in $s$), and let
\begin{equation}
\gamma  = \frac{M L}{N}
\label{eq:gamma}
\end{equation}
denote the density of $x$ in $y$ (that is, the fraction of entries of $y$ occupied by occurrences of $x$.) The separation condition %~\eqref{eq:spacing} 
imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

@@

We will denote by $\M_l$ the moments of $X$:
%
\begin{align}
%
\M_1[i] = \E X[i], \quad 0 \le i \le L-1,
%
\end{align}
%
\begin{align}
%
\M_2[i,j] = \E X[i] X[j], \quad 0 \le i,j \le L-1,
%
\end{align}
%
and
%
\begin{align}
%
\M_3[i,j,k] = \E X[i] X[j] X[k], \quad 0 \le i,j,k \le L-1.
%
\end{align}

We will also denote by $\L_l$ the autocorrelations of $X$, $l = 1,2,3$:
%
\begin{align}
%
\L_1 = \sum_{i=0}^{L-1} \M_1[i] \in \R,
%
\end{align}
%
\begin{align}
%
\L_2(\Delta) = \sum_{i=0}^{L-1} \M_2[i,i+\Delta]  \in \R^L,
%
\end{align}
%
and
%
\begin{align}
%
\L_3(\Delta_1,\Delta_2) = \sum_{i=0}^{L-1} \M_3[i,i+\Delta_1,i+\Delta_2]  \in \R^{L \times L}.
%
\end{align}

Note that in the strongly-separated model, the first three observed moments are, respectively, $\L_1$, $\L_2(\Delta)$, and $\L_3(\Delta_1,\Delta_2)$.

We will prove the following result:

\begin{proposition} \label{prop:poisson}
	The first moment of the Poisson process is equal to:
	%
	\begin{align}
	\gamma \L_1.
	\end{align}
	%
	The second moment vector of the Poisson process has entries
	%
	\begin{align}
	(\gamma \L_1)^2 + \gamma \L_2(\Delta),
	\end{align}
	%
	for $0 \le \Delta < L$.

	The third moment matrix of the Poisson process has entries
	\begin{align}
	(\gamma \L_1)^3 + \gamma \L_1  \cdot ( \gamma\L_2(\Delta_1) + \gamma\L_2(\Delta_2) + \gamma\L_2(\Delta_2-\Delta_1)) + \gamma \L_3(\Delta_1,\Delta_2) ,
	\end{align}
	for $0 \le \Delta_1,\Delta_2 < L$.
\end{proposition}

@@

\begin{rem}

	The expected values of the non-zero (for second moment) and off-diagonal (for third moment) terms are the same as without noise, as is true for the strongly-separated case. The same proof of almost sure convergence from my notes for the strongly-separated case also goes through verbatim.

\end{rem}
